---
title: "ELECTORAL BUDGET 2022-KENYA"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#d1e0e0"
      fg: "#000000" 
      primary: "#669900"
      base_font:
        google: Roboto
      code_font:
        google: Ubuntu
    orientation: columns
    vertical_layout: fill
    social: menu
    source_code: embed
    runtime: shiny
---

```{r setup, include=FALSE}
#import libraries
library(flexdashboard)
library(echarts4r)
library(dplyr)
#import data
data <- readxl::read_xlsx('data/data.xlsx')
```

Column {data-width=500}
-----------------------------------------------------------------------
### Statistics

**Key**

* Registered voters - 22,162,362
* Budget - $383,560,000 Million
* Per Registered voter - $17
* Elective positions - 6

### Elective positions

``` {r}
renderTable(
data.table::data.table(Position = c('President','Governor','Senator','Women Representative','Member of National Assembly','Member of County Assembly','Total'),
                      `Total Count` = c('1','47','47','47','290','1450','1882')
                      )
)

```

Column {data-width=500 .tabset}
-----------------------------------------------------------------------
### Comparative Data

``` {r}
renderTable (
  data.table::data.table(Country = c(data$Country),
                         Year = c(as.character(data$Year)),
                         Expenditure = c(paste('$',as.character(data$Expenditure)))
                      )
)
```
 

### Graph

```{r}
#radial gradient background
radial_gradient <- htmlwidgets::JS(
  "new echarts.graphic.RadialGradient(
    0.2, 0.7, 0.9,
    [
     {offset: 0, color: '#f7f8fa'},   
    { offset: 1, color: '#cdd0d5'}
    ])")
 # plot
data |>
  arrange(desc(Expenditure)) |> #arrange in descending order
  e_charts(Country) |> #x-axis
  e_bar(Expenditure, # y-axis
        bind = Year,
        name='Country',
        showBackground=TRUE,
        emphasis = list(
          focus = "item"),
        itemStyle=list(
          color = htmlwidgets::JS("
          function(params) {
          var colorList = ['#99e600','#00e6e6','#00e6e6','#00e6e6','#00e6e6','#00e6e6','#00e6e6'];
          return colorList[params.dataIndex]
          } 
          "),
          shadowBlur=1, #customise bars
          shadowColor='#4f1721',
          shadowOffsetY=2)
        )|>
  e_animation(duration = 4000)|> #duration in plot animation
  e_tooltip(trigger='item',
            backgroundColor='#ffffcc',
            formatter =
              htmlwidgets::JS("
                              function(params){
                              return('<strong>' + params.value[1] +
                              '</strong><br />Year: '  + params.name +  
                              '<br />Expenditure: ' +'</> $' +
                              params.value[0]) 
                              }
                              "))|>
  e_axis_labels(x='Countries',y = 'Expenditure') |> #axis labels
  e_legend(show = FALSE) |>
  e_title(text=paste('Expenditure per Registered Voter'),
          subtext="Analyst: Jefferson Ndeke  Insights: BBC News Africa",link = 'https://www.facebook.com/watch/?v=797291747927851',
          left='center',top=10,
          sublink = 'https://github.com/ndeke254') |> #title
  e_toolbox_feature(feature = c("restore","saveAsImage"),type='bar') |> #features to interact
  e_datazoom(type='inside') |> #zoom to view
  e_grid(show=TRUE,backgroundColor=radial_gradient, 
         containLabel=TRUE,left='3%',right='7%',bottom='3%')|> #grid
  e_color(background='#cdd0d5',color = c("green", "red")) |>
  e_x_axis(axisTick=list(
    alignWithLabel= TRUE
  ),
  splitLine=list(
    lineStyle=list(
      type='dashed'))) |>
  e_y_axis(scale=TRUE,
           axisTick=list(show=TRUE,
             alignWithLabel= TRUE
             ),
           splitLine=list(
             lineStyle=list(
               type='dashed')),
           formatter = e_axis_formatter('currency')) |>
  e_flip_coords() 

```

### Cost Analysis

**KENYA**

* Kenya's General Election 2022 was among the most expensive in the world
* Apart from Uganda and Zambia, all the other countries are richer than Kenya in terms of GDP
 
 **But why?**
 
 * A high number of elective positions
 * Lack of infrastructure like network coverage in some poling stations
 * Lack of trust - Import voting materials like ballot papers and voting kits
 
 *However compared to 2017, this year's budget is significantly lower due to the already existing technology.*
 
 **My take**
 
*Kenya, a developing country, has a large wage bill and high budget on recurrent expenditure. This is a call for a reduction on overall representation of the tax payer.*
